From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam

### 2025-03-03 manuscript.qmd
Update LLM results

As previously mentioned, the web scraping process also captured unstructured information from the property descriptions, which often contained valuable details not captured in the primary fields. 
In this subsection, we detail the data cleaning process using a pre-trained large language model (LLM) to extract structured information from the unstructured text. 
The LLM used was the cloud-based OpenAI o1-mini reasoning model[^1], accessed using the `{ellmer}` R package [@wickham2025ellmer] via an API.
<!-- The LLM used was a DeepSeek R1 distilled Qwen 14B [@deepseek2025deepseekr1;@qwen2025qwen25], accessed using the `{ellmer}` R package [@wickham2025ellmer] via the Ollama[^2] API, a local interface platform to the LLM. -->

[^1]: <https://openai.com/index/openai-o1-mini-advancing-cost-efficient-reasoning/>
[^2]: <https://github.com/ollama/ollama>

@@

The verbose descriptions were fed into the model one at a time using a loop, with the LLM extracting and returning the relevant details. 
Note that this loop was not parallelised, as the LLM already utilises multiple cores, and further parallelisation would not yield significant efficiency gains due to resource constraints.
The cleaned results were then parsed and stored in a data frame, which was then subjected to manual data-type validation to ensure conformity with the existing data set. 
We processed 5,429 descriptions in total from 2020 to 2025 using this method.
It takes, on average, 7.51 seconds to process a single description.

To test the accuracy of the LLM data extraction, a test data set of 100 artificially generated descriptions was created, with each description written in the style of a typical property listing.
The test data also contained missing information in selected property characteristics, which the LLM was expected to handle correctly.
For each house characteristic, the accuracy of the LLM was calculated by comparing the extracted value to the ground truth.
For numeric variables, values were considered accurate if they fell within 2% of the corresponding ground truth value. 
For character variables, accuracy was determined using the normalized Levenshtein distance, ensuring differences remained within a set threshold. 
Missing values were counted as correct if both the extracted and ground truth values were missing. 
Accuracy was then averaged both per characteristic and across all characteristics.

Our experiments show that the LLM data extraction process achieved an overall accuracy rate of 98.2%.
Errors mostly stemmed from incorrect status classification, most likely due to the vagueness of the descriptions.
In fact, excluding the `status` variable and the `location` variable, which will be processed further, the accuracy rate increases to 99.1%.

In conclusion, we found the LLM to be a valuable tool for extracting structured information from unstructured text, significantly reducing the time and effort required for data cleaning.
Large errors were still able to be corrected manually, by filtering for outliers (values exceeding three standard deviations from the mean in magnitude) or inconsistencies in the variables, which may be a result of model hallucinations.
We note that the data cleaning procedure may also be run on local hardware, for example using the Ollama[^2] API, a local interface platform to the LLM.
See the appendix for further details.

@@

Despite efforts to ensure data quality, some limitations remain. Duplicate listings were carefully reviewed and removed, though there remains a slight possibility that duplicated records still exist in the data set. 
The spatial data coverage is also heavily skewed towards the Brunei-Muara district, which accounts for `{r} scales::percent(prop.table(table(hsp$district))[2], accuracy = 0.1)` of the listings. 
This is a reflection of the district’s higher population and greater volume of property transactions [@jamil2024spatiotemporal], which may bias analyses toward this region. 
Furthermore, while we have confidence in the data quality from 2015 to 2024, property price trends between 1993 and 2014 cannot be fully verified. Nonetheless, this study serves as a valuable starting point. 
Future research could benefit greatly from access to administrative transaction data, which would allow for more comprehensive and accurate analyses.

>>>>

As previously mentioned, the web scraping process also captured unstructured information from the property descriptions, which often contained valuable details not captured in the primary fields. 
In this subsection, we detail the data cleaning process using a pre-trained large language model (LLM) to extract structured information from the unstructured text. 
<!-- The LLM used was the cloud-based OpenAI o1-mini reasoning model[^1], accessed using the `{ellmer}` R package [@wickham2025ellmer] via an API. -->
The LLM used was a DeepSeek R1 distilled Qwen 14B [@deepseek2025deepseekr1;@qwen2025qwen25], accessed using the `{ellmer}` R package [@wickham2025ellmer] via the Ollama[^2] API, a local interface platform to the LLM.

<!-- [^1]: <https://openai.com/index/openai-o1-mini-advancing-cost-efficient-reasoning/> -->
[^2]: <https://github.com/ollama/ollama>

@@

The verbose descriptions were fed into the model one at a time using a loop, with the LLM extracting and returning the relevant details. 
Note that this loop was not parallelised, as the LLM already utilises multiple cores, and further parallelisation would not yield significant efficiency gains due to resource constraints.
The cleaned results were then parsed and stored in a data frame, which was then subjected to manual data-type validation to ensure conformity with the existing data set. 
We processed 5,429 descriptions in total from 2020 to 2025 using this method, with a run time of approximately 20 hours spread over 3 machines.
It takes, on average, 150 seconds to process a single description.

To test the accuracy of the LLM data extraction, a test data set of 100 artificially generated descriptions was created, with each description written in the style of a typical property listing.
The test data also contained missing information in selected property characteristics, which the LLM was expected to handle correctly.
For each house characteristic, the accuracy of the LLM was calculated by comparing the extracted value to the ground truth.
For numeric variables, values were considered accurate if they fell within 2% of the corresponding ground truth value. 
For character variables, accuracy was determined using the normalized Levenshtein distance, ensuring differences remained within a set threshold. 
Missing values were counted as correct if both the extracted and ground truth values were missing. 
Accuracy was then averaged both per characteristic and across all characteristics.
For further details, please consult the appendix.

Our experiments show that the LLM data extraction process with the deepseek-r1:14b model achieved an overall accuracy rate of 96.9%.
Errors mostly stemmed from incorrect status classification, most likely due to the vagueness of the descriptions.
<!-- In fact, excluding the `status` variable and the `location` variable, which will be processed further, the accuracy rate increases to 97.3%. -->
As for the rest of the variables, obvious errors were able to be corrected manually, by filtering for outliers (values exceeding three standard deviations from the mean in magnitude) or inconsistencies in the variables, which may be a result of model hallucinations.
Overall, the LLM was found to be a valuable tool for extracting structured information from unstructured text, significantly reducing the time and effort required for data cleaning. 
<!-- Users may wish to exclude these records from their analysis if they are concerned about the accuracy of the extracted data. -->

@@

Despite efforts to ensure data quality, some limitations remain. Duplicate listings were carefully reviewed and removed, though there remains a slight possibility that duplicated records still exist in the data set. 
The spatial data coverage is also heavily skewed towards the Brunei-Muara district, which accounts for `{r} scales::percent(prop.table(table(hsp$district))[2], accuracy = 0.1)` of the listings. 
This is a reflection of the district’s higher population and greater volume of property transactions [@jamil2025spatiotemporal], which may bias analyses toward this region. 
Furthermore, while we have confidence in the data quality from 2015 to 2024, property price trends between 1993 and 2014 cannot be fully verified. Nonetheless, this study serves as a valuable starting point. 
Future research could benefit greatly from access to administrative transaction data, which would allow for more comprehensive and accurate analyses.
